---
title: "New York Times Best Sellers"
author: "Ben Rubin"
date: 2023-02-02
output: html_output
execute: 
  echo: true
  warning: false
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(rbin)

```

```{r Load, include=FALSE}

tt <- tt_load("2022-05-10")

```

```{r Readme, eval = interactive(),include=FALSE}

tt

```

# 

The New York times has ranked many books in the past century. Here I have analysed which books were ranked #1 for the longest time, and which authors had the most NYT best sellers.

This work was inspired by the article [Finding Trends in NY Times Best Sellers](https://towardsdatascience.com/finding-trends-in-ny-times-best-sellers-55cdd891c8aa). The data set can be found [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-05-10).

# Data Wrangling

```{r Wrangle}

nyt_full <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-10/nyt_full.tsv')
nyt_titles <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-10/nyt_titles.tsv')



rank_1=nyt_full %>% filter(rank==1) %>% group_by(title,title_id) %>%  count()
book_year=nyt_titles %>% summarize(year,author,id)
rank_1=left_join(rank_1,book_year, by=c('title_id'='id')) %>% 
  mutate(decade = cut(year, breaks=c(1930,1940,1950,1960,1970,1980,1990,2000,2010,2020,2030),labels = c("1930", "1940","1950","1960","1970","1980","1990","2000","2010","2020"), right=FALSE)) %>% arrange(desc(n),desc(decade))

rank_2=rank_1 %>% filter(title_id!=481,title_id!=2026,title_id!=6157,title_id!=2126,title_id!=6366,title_id!=2816,title_id!=4897,title_id!=4723,title_id!=4918,title_id!=7239,title_id!=414)
#For each decade I want all books, but want to highlight the books that were ranked number 1 the longest. bin the decades

top_ranked=rank_1 %>% group_by(decade) %>% top_n(1,n) %>% arrange(decade)

rank_3=nyt_full  %>% group_by(author) %>% distinct(title) %>% filter(author!="James Patterson and Maxine Paetro") %>% count() %>% filter(n>36)
```

# Analysis

Of the Novels ranked by the NYT, <u>Dan Brown's "The Davinci Code</u>, ranked at #1 the longest for 59 weeks. Followed by Delia Owen's <u>Where the Crawdads Sing</u> at 54 weeks, and James A. Michener's <u>Hawaii</u> at 49 weeks.

```{r Visualize, results=T}
rank_2 %>% 
  ggplot(aes(x=decade,y=n,color=decade))+
  geom_jitter(alpha=0.5)+
  geom_jitter(data=top_ranked, inherit.aes = T, aes( x=decade, y= n),color="black", height = 0)+
  theme_minimal()+
  scale_y_continuous(limits = c(0,80))+
  labs(title = "The Decade's Longest #1 Ranked NYT Best Sellers", x="Decade",
      y="Weeks ranked as #1")+
  annotate("text", x = top_ranked$decade[1], y = top_ranked$n[1], label = 
             "ANTHONY \n ADVERSE", vjust = -3, fontface = "bold", size=2.5) +
  annotate("text", x = top_ranked$decade[2], y = top_ranked$n[2], label = 
           "GONE WITH \n THE WIND", vjust = -1, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[3], y = top_ranked$n[3], label = 
           "THE \n ROBE", vjust = -1, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[4], y = top_ranked$n[4], label = 
             top_ranked$title[4], vjust = -1, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[5], y = top_ranked$n[5], label = 
             "THE \n SOURCE", vjust = -1, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[6], y = top_ranked$n[6], label = 
             "LOVE \n STORY", vjust = -1, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[7], y = top_ranked$n[7], label = 
             "THE \n COVENANT", vjust = -1, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[8], y = top_ranked$n[8], label = 
             "THE BRIDGES \n OF \n MADISON COUNTY", vjust = -1, fontface = "bold", 
           size=2.5)+
  annotate("text", x = top_ranked$decade[9], y = top_ranked$n[9], label = 
             "THE \n DA VINCI \n CODE", vjust = -1, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[10], y = top_ranked$n[10], label = 
             "WHERE THE \n CRAWDADS SING",vjust = -1,hjust=0, fontface = "bold", size=2.5)+
  annotate("text", x = top_ranked$decade[11], y = top_ranked$n[11], label = 
             "AMERICAN \n DIRT", vjust = -1, fontface = "bold", size=2.5)+
  theme(plot.title = element_text(vjust = 1), plot.subtitle = 
        element_text(vjust = 2), 
        axis.title.y = element_text(face = "bold")) +
  theme(legend.position = "none")



  
```

# 

Among the many authors on the list, Danielle steele has the most #1 best sellers at 116 books. Followed by Stuart Woods with 63, and Stephen King with 54.

```{r}
ggplot(rank_3,aes(x=reorder(author,n), y=n, fill=author))+
  geom_bar(stat = 'identity')+
  theme_minimal()+
  coord_flip()+
  labs(title ="Authors with the most NYT Best Sellers", x="Authors", y='Books')+
  theme(plot.title = element_text(vjust = 1), plot.subtitle = 
        element_text(vjust = 2), 
        axis.title.y = element_text(face = "bold")) +
  theme(legend.position = "none")
```
